\documentclass[12pt]{article}
\input{../icmp-macros.tex}
\setheadings{MTH 493/593 --- Introduction to Computational Mathematics --- Project 1}

\begin{document}
\sffamily

\section{Submission Instructions}
The project must be submitted through BBLearn as a Jupyter notebook. Create a new notebook in Google Colab. Please name your notebook as follows:

\begin{center}
Project 1 - Your Name - Your CSU ID
\end{center}

For example, Alice Mathperson with ID 1234567 would name her notebook:

\begin{center}
Project 1 - Alice Mathperson - 1234567
\end{center}

You are allowed to make multiple submissions in BBLearn. However, only the final submission will be graded. In your final submission, make sure of the following:

\begin{itemize}
\item The notebook must run without errors. Before submitting the notebook, check that all cells execute correctly, \emph{in the order they appear in the notebook}.
\item There is one exception to the rule above. Sometimes it is useful that we know, or suspect, to produce errors. In this case, provide an explanation in a text cell for the error.
\item Add text cells to explain what you are doing in each step, and state your conclusions. Your project should be readable as a mathematical paper.
\item Your code should not only be correct, but well structured and clear. Use meaningful variable names and add comments where you feel necessary.
\end{itemize}

\section{Project Statement}
Let's suppose that we need to repeatedly solve quadratic equations, for a large number of cases. In this project, you will define a function \texttt{quadratic\_equation} that outputs the solutions to a the quadratic equation $ax^2+bx+c=0$ given the coefficients $a$, $b$ $c$. We start with a reasonably simple implementation, and then add features to handle more complicated cases. At the end, we will have professional quality code to solve quadratic equations.

\subsection{Part 1 --- Define and test your function}
Write a function that implements the quadratic formula:
\[
x_1,x_2=\frac{-b\pm \sqrt{b^2-4ac}}{2a}
\]
Your code will have the following structure:

\begin{lstlisting}
def quadratic_equation(a, b, c):
    ... formula breakdown ...
    x1 = ...
    x2 = ...
    return x1, x2
\end{lstlisting}

It is recommended that long formulas are avoided, by doing the computation in a series of statements. For example, the function can start by computing the discriminant $b^2-4ac$, and then using this partial computation in the evaluation of the solutions.

Test your code with at least ten test cases. Each test case consists of values for the coefficients $a$, $b$, $c$. Make sure to include tests for the following cases:

\begin{itemize}
\item The solutions are real and distinct.
\item The solutions are not real.
\item The two solutions are identical.
\item The coefficients are complex numbers.
\item The coefficient $a$ is zero.
\end{itemize}

Make notes of any errors that occur.

\subsection{Part 2 --- Handling Complex Solutions}

We now want to polish our function in a way that it produces correct and meaningful results in all cases. We want a function that works in any case, and returns the right type of solution, specially regarding the issue of complex coefficients and/or solutions.

Use the following structure for the function definition:

\begin{lstlisting}
def quadratic_equation(a, b, c):
    if type(a) == complex or type(b) == complex or type(c) == complex:
        ... computation in the case of complex coefficients ...
        x1 = ...
        x2 = ...
    else:
        delta = b ** 2 - 4 * a * c
        if delta >= 0:
            ...  real coefficients, but complex roots ...
            x1 = ...
            x2 = ...
        else:
            ... real coefficients, real roots ...
            x1 = ...
            x2 = ...
return x1, x2
\end{lstlisting}

Test the new version of the function with the same test cases you used in Part 1.

\subsection{Part 3 --- Handling Exceptions}

At this point, there is still one case that your function does not handle well: if the coefficient $a$ is zero. To take care of this case, we will make the function signal an error with a meaningful message. Add the following code right after the \lstinline{def quadratic_equation(a, b,c)}:

\begin{lstlisting}
if a == 0:
    raise ValueError("coefficient a cannot be zero")
\end{lstlisting}

Repeat all tests with the new version of the function.
\end{document}
